### 2019-06-25 size_differentiated_export.tex
Updates from Overleaf

%\documentclass[]{article}

%\setlength{\parskip}{.5cm}
%\usepackage{natbib}
%\usepackage[table,xcdraw]{xcolor}
%\usepackage{graphicx}
%\usepackage{hyperref}
%\usepackage{amsmath}
%\usepackage{appendix}
%\usepackage[left=3cm,right=3cm]{geometry}
  % Apply watermark to document
%\usepackage{draftwatermark}
%\SetWatermarkLightness{0.9}
%\SetWatermarkScale{15}
%\SetWatermarkText{DRAFT}

\documentclass[draft,linenumbers]{agujournal2018}
\usepackage{apacite}
\usepackage{url} %this package should fix any errors with URLs in refs.

% FIXME command
\newcommand{\fixme}[1]{\color{red}$<$\textbf{FIX ME: #1}$>$\color{black}}

% Trackchanges package
%\usepackage{trackchanges}
%% PLEASE USE THE FOLLOWING COMMANDS
%   \note[editor]{The note} 
%   \annote[editor]{Text to annotate}{The note} 
%   \add[editor]{Text to add} 
%   \remove[editor]{Text to remove} 
%   \change[editor]{Text to remove}{Text to add}

%\draftfalse

\journalname{Global Biochemical Cycles}

\begin{document}

%% ------------------------------------------------------------------------ %%
%
%  TITLE
%
%% ------------------------------------------------------------------------ %%
\title{Size-differentiated Export in different Dynamical Regimes in the Ocean}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%
\authors{M. Dever\affil{1}, D. Nicholson\affil{1}, M. M. Omand\affil{2}, and A. Mahadevan\affil{1}}
\affiliation{1}{Woods Hole Oceanographic Institution, Woods Hole, MA 02543, USA}
\affiliation{2}{Graduate School of Oceanography, University of Rhode Island, Narragansett, Rhode Island, USA}

\correspondingauthor{M. Dever}{mdever@whoi.edu}
%% ------------------------------------------------------------------------ %%
%
%  KEYPOINTS
%
%% ------------------------------------------------------------------------ %%

%\begin{keypoints}
%\item = enter point 1 here =
%\item = enter point 2 here =
%\item = enter point 3 here =
%\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
%% ------------------------------------------------------------------------ %%
\begin{abstract}
% Mat: I think that even with the submesoscale mechanism, export is mainly driven by gravitational sinking. So I'm moving the "Traditionally" to later. 
Export of particulate organic carbon (POC) is  mainly driven by gravitational sinking. Thus, traditionally, it is thought that larger, faster-sinking particles make up most of the POC export flux. However, this need not be the case in a dynamic oceanic flow field, where the ocean velocity can influence the descent rate of particles. Particles with different settling speeds are released in two process-oriented model simulations of an upper ocean eddying flow to evaluate the impact of (1) the ocean dynamics on the respective contribution of the different sinking-velocity classes to POC export, and (2) the particle number size-spectrum slope. The analysis reveals that the leading export mechanism changes from gravitationally-driven to advectively-driven as submesoscale dynamics become more important. The vertical velocity associated with submesoscale dynamics enhances the contribution of slower-sinking particles to POC export. A steeper particle size spectrum, also increases the contribution of smaller, slower-sinking particles. Implementing a remineralization scheme generally decreases the total amount of biomass exported, but its impact is weaker in dynamical regimes where submesoscale dynamics are present and export is advectively-driven. Under specific conditions, remineralization processes counter-intuitively enhance the role of slower-sinking particles to the point where these slower-sinking velocity classes dominate the export, therefore challenging the traditional paradigm for POC export. This study demonstrates that slow-sinking particles can be a significant contribution, and at times, even dominate the export flux. 
\end{abstract}

%% ------------------------------------------------------------------------ %%
%
%  BEGIN ARTICLE
%
%% ------------------------------------------------------------------------ %%
%\begin{article}

%\maketitle

%{\color{red} NOTE: Following Roo's suggestion, I looked at the spectrum of biomass per velocity class vs time to see how it evolves. As Melissa suggested, it will be cool to include into a presentation, but I don't think it is worth including in the paper. I believe it would introduce too much additional concepts that would dilute the conclusions: the spectral shape is mostly determine by the pulse of particle crossing the export horizon. A constant reseeding of particles would be necessary to clearly put our point across using biomass spectra.}

\section{Introduction}

A major pathway transferring carbon from the atmosphere to the ocean is through photosynthesis and the production of Particulate Organic Carbon (POC) in the upper ocean. Sinking POC then allows the transfer of organic carbon from the upper ocean to the interior, leading to carbon sequestration over long time scales \citep{Falkowski_1998}. Understanding the mechanisms driving the export of POC from the ocean's surface to the interior is therefore crucial to better constrain Earth's carbon budget.

Traditionally, POC export is thought to be mainly driven by larger, faster-sinking, particles capable of gravitationally-sinking below the export depth over timescales shorter than remineralization timescales. This is a reasonable assumption if one considers that vertical advective velocities in the ocean are generally much weaker than velocities associated with gravitational sinking. However, recent work demonstrates that the presence of mesoscale eddies can enhance carbon export through two main processes: First, by increasing primary productivity through the injection of nutrients into the euphotic layer, therefore increasing the particle concentration, logically leading to a greater export \citep{Levy_1998}. Second, by generating larger vertical velocities along the front located at the eddy perimeter \citep{vanHaren_2006, Waite_2016}. These studies demonstrate that ocean dynamics potentially play an important role in driving the carbon export from the euphotic layer to the ocean interior. This conclusion is further supported by a recent, yet large, body of literature focusing on ocean dynamics at submeso-scales (1-10 km) and the impact on biogeochemical processes \citep[see reviews in][]{Thomas_2008,Klein_2009,Mahadevan_2016}. Submesoscale dynamics, characterized by Rossby numbers of order 1, typically develop in filaments in areas where sharp density fronts exist \citep{McWilliams_2016}. In this dynamical regime, geostrophic balance fails and a secondary ageostrophic circulation develops across the front, capable of generating large vertical velocities on the order of 100 m/day \citep{FoxKemper_2008,Mahadevan_2016}. On the denser side of the front, the vorticity is cyclonic and associated with downwelling velocities, while anticyclonic vorticity and upwelling is expected on the lighter side of the front. The distribution of relative vorticity across a front is asymmetric and skewed toward cyclonic vorticity \citep{Rudnick_2001}, leading to more localized and more intense downwelling regions, as opposed to weaker and larger upwelling regions \citep{Mahadevan_2006}. Just like at meso-scales, enhanced vertical velocities can generate either a local bloom by supplying nutrients to the sunlit layer of the ocean \citep{Mahadevan_2012}, or can significantly increase the export of POC to the ocean interior \citep{Levy_2012, Estapa_2015,Omand_2015}.

The downwelling velocities $\mathcal{O}$(100 m/day) generated at submeso-scales provide a physical mechanism capable of competing with gravitational sinking and thus exporting particles over a larger portion of the particle size spectrum. Through this mechanism, smaller particles can efficiently be exported over timescales shorter than their remineralization timescales, despite their slower sinking velocities. Considering the fact that, although containing less biomass per particle, smaller particles are far more numerous than larger particles, the impact of submesoscale dynamics on the export of POC is potentially significant. Submesoscale processes are, by definition, associated with spatial scales $\mathcal{O}$(1 km) and time scales $\mathcal{O}$(1 day), which makes it difficult to observe in-situ. Additionally, and despite recent methodological progress \citep{McDonnell_2010}, direct measurements of particles sinking velocities are difficult to obtain, and are more often than not inferred from key parameters such as particle type, size, and density.

In this study, we therefore rely on a submesoscale-resolving non-hydrostatic model coupled with a particle-tracking module to explore the necessary conditions for slow-sinking particles to significantly contribute to carbon export, as a function of (1) the dynamics of the flow field, (2) the slope of the sinking velocity spectrum, and (3) the remineralization timescale. Particles in the model are prescribed with both a constant and time-varying sinking velocity to mimic a remineralizing behavior. Particles are released in two fundamentally different flow fields in terms of dynamics based on observed conditions in the Northeast Pacific in the summer, where ocean dynamics are characterized by low Rossby numbers and weak vertical advective velocities, and in the winter, where ocean dynamics include submesoscale frontal structures and local Rossby numbers $\mathcal{O}$(1). Both simulations and the particle-tracking module are described in Section \ref{sec: Methods}. The impact of particles characteristics and ocean dynamics on the export of POC is quantified in Section \ref{sec: Results}, and discussed in Section \ref{sec: Discussion}. Section \ref{sec: Conclusions} summarizes the key conclusions of the study.

\section{Methods}
\label{sec: Methods}
\subsection{Model setup and domain}

This study uses a non-hydrostatic, three-dimensional, numerical model called PSOM \citep{Mahadevan_1996a,Mahadevan_1996b}. To simulate an eddy field that is representative of the Northeast Pacific Ocean, the model domain is set in a channel configuration with periodic east-west boundaries, and solid boundaries in the south and north. The domain covers 112 km in the x-direction, 304 km in the y-direction, and 1000 m in the vertical (Figure \ref{fig: model_domain}). The horizontal resolution is 500 m, while a stretched grid is used in the vertical with 32 levels ranging in thickness from 1.25 m near the surface to 70 m at the lowermost level. The model is integrated numerically in time and evolves the temperature, salinity, free-surface height, pressure, and three-dimensional velocity field from an initial state, subject to momentum and buoyancy fluxes applied through the surface boundary conditions.

% FIG1
\begin{figure}[ht]
	\centering
	\includegraphics[width = .6\linewidth]{figures/Fig1_model_domain}
	\caption{PSOM model setup. (a) Meridional profile of zonal wind scaling coefficient that multiplies $\tau_x$, with a tapering at the north and south boundaries to reduce upwelling and downwelling, respectively. (b) Restoration factor (color shading) used to dampen internal wave reflection at boundaries as well as curl-driven upwelling and downwelling, along with constant density contours showing the three fronts (black line). (c) Surface net heat flux anomaly superimposed to the heat fluxes prescribed at $y = 150$ km.}
	\label{fig: model_domain}
\end{figure}

Time-varying wind stress and heat flux are prescribed at the surface boundary. Time series are computed from measurements collected at Station Papa and available through the Pacific Marine Environmental Laboratory. Daily wind stress and net heat fluxes are calculated over the period 2007-2016 to produce a year-long climatology. A squared low-pass filter with a cut-off frequency of 8.5 days is applied to both time series to remove high-frequency variability. In all numerical experiments, the simulations are first allowed to spin-down for the first 5 days with no forcing applied at the surface boundary. Surface wind stress and heat fluxes are then linearly ramped up between days 5 and 10 of the simulation, to reach realistic values at day 10. % (Figures \ref{fig: Papa_summer} and \ref{fig: Papa_winter}).
While the meriodional component, $\tau_y$, is set to zero, the zonal component of the wind stress, $\tau_x$, is prescribed at the surface throughout the model domain and is tapered at the northern and southern boundaries to avoid excessive Ekman-driven upwelling and downwelling %that could contaminate the middle third of the model domain, where the analysis is focused 
(Figure \ref{fig: model_domain}a). A restoration timescale is prescribed to contain the curl-driven upwelling and downwelling regions generated by the tapering of the wind stress, as well as to limit internal wave reflection at the solid boundaries back into the domain (Figure \ref{fig: model_domain}b). While net surface heat fluxes are homogeneous in the zonal direction, a meridional gradient is maintained throughout the simulation. The meridional gradient was determined from the North American Regional Reanalysis (NARR) product \citep{Mesinger_2006}, and set to 1/24 W/m$^2$/km (Figure \ref{fig: model_domain}c).

Initial hydrographic conditions are determined from a three-dimensional gridded field of temperature and salinity from Argo floats \citep{Gaillard_2009,Gaillard_2016}. Argo data is averaged monthly over the period 2002-2012 and two different months are used to initialize the two main numerical experiments for this study: Climatological conditions in April are used to initialize the \textit{Papa\_summer} experiment, while January climatological conditions are used to initialize the \textit{Papa\_winter} experiment (Table \ref{tab: meso_vs_submeso}). The north-south background density gradient is then intensified into three fronts located at $y = 75$, $y = 150$, and $y = 225$ km (Figure \ref{fig: model_domain}). The amplitude of the density gradient associated with the three fronts is determined from the probability distribution function (PDF) of the density gradients measured by underwater gliders deployed around Station Papa over the period 2008-2010 \citep{Pelland_2016}. To reduce model spin-up time, density fronts are perturbed by a sinusoidal wave with a wavelength close to the 1st baroclinic deformation radius ($\lambda = $ 66 km). Similar PSOM configurations were successfully used in previous studies \citep{Mahadevan_2012,Omand_2015}.

Two main experiments are conducted using the same configuration of PSOM, where only initial conditions and surface forcings are varied: \textit{Papa\_summer} aims at generating ocean dynamics representing conditions in the Northeast Pacific in the summertime. Summer ocean dynamics are characterized by a flow generally in geostrophic balance, with relatively weak density gradients and low Rossby numbers ($\ll$1). \textit{Papa\_winter} aims at capturing wintertime ocean conditions in the region. A different dynamical regime is expected to dominate during wintertime when mixed layers are deeper and lateral density gradients enhanced, with sharper density fronts, filament-like features and localized Rossby number $\mathcal{O}$(1) over spatial scales $\mathcal{O}$(1 km) \citep{Mensa_2013,Callies_2015,Thompson_2016}. The individual characteristics of each of \textit{Papa\_summer} and \textit{Papa\_winter} are detailed below.

\subsubsection{\textit{Papa\_summer} Model Experiment}

In \textit{Papa\_summer}, PSOM is initialized based on climatological Argo data in April. The magnitude of the density gradient across the front is set to 3.34$\times$10$^{-6}$ kg/m$^3$/m, which corresponds to the 95$^{th}$ percentile of the PDF of density gradients measured in April from glider data collected in the region (Figure \ref{fig: Papa_summer} and Table \ref{tab: meso_vs_submeso}). The model is run with a timestep of 216 s and is allowed to spin-up for 60 days, allowing summer stratification to develop. The model is then run for 30 additional days, saving instantaneous model fields every 3 hours for particle tracking. The month of April is chosen for initialization so the experiment would capture the onset of positive net heat fluxes, and the summer restratification that ensues in July-August (Figure \ref{fig: Papa_summer}). In this region, the summer stratification is associated with large primary productivity, particle production, and POC export \citep[e.g., fecal pellets, dead phytoplankter; ][]{Plant_2016}.

%TAB1
\begin{table}[t]
\caption{Summary of the key characteristics of PSOM experiments \textit{Papa\_summer} and \textit{Papa\_winter}.}
	\label{tab: meso_vs_submeso}
	\centering
 \centering
\begin{tabular}{|l|c|c|}
		\hline
											& \textit{Papa\_summer} & \textit{Papa\_winter} \\
		\hline
		Time period 								& April -- July             & January -- March          \\
		Spin-up								        & 60 days                   & 50 days                   \\
		Advective timestep				            & 216 s                     & 108 s                     \\
		Horizontal diffusivity			            & 1 m$^2$ s$^{-1}$          & 0.2 m$^2$ s$^{-1}$        \\
		Restoration timescale			            & 3 days                    & 15 days                   \\
		Zonal wind stress                	        & 0 -- +0.16 N m$^{-2}$      & -0.05 -- +0.17 N m$^{-2}$  \\	
		Surface heat flux               	        & -46.8 -- +167.5 W m$^{-2}$ & -57.6 -- +15.3 W m$^{-2}$  \\
		Maximum M$^2$ ($\times$10$^{-8}$)	        &                           &                           \\[-.25cm]
		\multicolumn{1}{|r|}{initial}	            & 3.2 s$^{-2}$              & 33.9 s$^{-2}$             \\[-.25cm]
		\multicolumn{1}{|r|}{spun-up}	            & 12.0 s$^{-2}$             & 50.0 s$^{-2}$             \\	
		Maximum N$^2$ ($\times$10$^{-4}$)	        &                           &                           \\[-.25cm]
		\multicolumn{1}{|r|}{initial}	            & 1.5 s$^{-2}$              & 1.6 s$^{-2}$              \\[-.25cm]
		\multicolumn{1}{|r|}{spun-up}	            & 3.1 s$^{-2}$              & 1.1 s$^{-2}$              \\	
		Averaged mixed layer depth                  &                           &                           \\[-.25cm]
		\multicolumn{1}{|r|}{initial}	            & 73 m                      & 85 m                      \\[-.25cm]
		\multicolumn{1}{|r|}{spun-up}	            & 11 m                      & 93 m                      \\	
		\hline
	\end{tabular}
\end{table}

% FIG2
\begin{figure}[t]
	\includegraphics[width = 1\linewidth]{figures/Fig2_Papa500_summer}
	\caption{PSOM configuration for \textit{Papa\_summer}. (a) Time series of net heat fluxes and wind stress prescribed at the surface. Notice the positive heat fluxes, as well as downfront winds (i.e. eastward) persisting throughout the experiment. (b)-(d) surface horizontal buoyancy gradients $M^2 = |\nabla_Hb|^2$ (in s$^{-2}$) at day of year (doy) 105, 135, and 165. Black contours show isopycnals (in kg/m$^3$; CI = 0.01 kg/m$^3$). (e) Vertical profile of the buoyancy frequency $N^2$ at day of year 105, 135, 165, and 195, showing the development of summer stratification centered at z = 30 m (solid lines). Monthly-average vertical stratification obtained from glider profiles collected in June and July are superimposed (dashed lines), along with the correlation coefficient between observations and model outputs.}
	\label{fig: Papa_summer}
\end{figure}

\subsubsection{\textit{Papa\_winter} Model Experiment}

In \textit{Papa\_winter}, PSOM is initialized based on climatological Argo data in January. The frontal gradient is set to 3.54$\times$10$^{-5}$ kg/m$^3$/m, which corresponds to the 99$^{th}$ percentile of the PDF of density gradients measured in January from glider data collected in the region (Figure \ref{fig: Papa_winter} and Table \ref{tab: meso_vs_submeso}). The model is allowed to spin-up for 50 days allowing for the prescribed fronts to become unstable. To accommodate for the larger density gradients and stronger velocities, the advective timestep is shortened to 108 s and the horizontal diffusivity is lowered to 0.2 m$^2$/s throughout the experiment. The model is run for 30 additional days, saving instantaneous model fields every 1.5 hours for particle tracking. The month of January is chosen for initialization so the experiment would capture the time of year where the mixed layer is the deepest, and Rossby number O(1) occur more frequently. The objective is for this experiment to contrast \textit{Papa\_summer} by capturing the statistics of ocean conditions dominated by submesoscale dynamics.

% FIG3
\begin{figure}[ht]
	\includegraphics[width = 1\linewidth]{figures/Fig3_Papa500_winter}
	\caption{PSOM configuration for \textit{Papa\_winter}. (a) time series of net heat fluxes and wind stress prescribed at the surface. Notice the mostly negative heat fluxes, as well as alternating zonal wind direction. (b)-(d) surface horizontal buoyancy gradients $M^2 = |\nabla_Hb|^2$ (in s$^{-2}$) at day of year (doy) 0, 30, and 50. Black contours show isopycnals (in kg/m$^3$; CI = 0.01 kg/m$^3$). (e) Vertical profile of the buoyancy frequency $N^2$ at doy 0, 30, 50, and 80, showing the persistence of the halocline between z = 80 and z =180 m throughout the experiment (solid lines). Monthly-average vertical stratification obtained from glider profiles collected in March is superimposed (dashed line), along with the correlation coefficient between observations and model outputs.}
	\label{fig: Papa_winter}
\end{figure}

\subsubsection{Validation}
\label{sec: glider_validation}

To ensure that PSOM simulations yielded realistic conditions for both \textit{Papa\_summer} and \textit{Papa\_winter}, distributions of horizontal ($M^2$) and vertical ($N^2$) buoyancy gradients are compared with glider observations collected over the period 2008-2009 \citep{Pelland_2016}.
During this period, underwater gliders sampled in a ``bow-tie" pattern centered on Station Papa. Gliders sample the water column in a triangular wave pattern, whose shape is easily affected by currents, due to the slow moving speed of the glider ($\sim$1 km/hr). It is therefore challenging to associate a specific spatial scale with gradients computed between glider profiles, as profile separation distances can be highly variable through depth and time. To circumvent this issue, horizontal buoyancy gradients are computed between each pair of glider profiles available within one branch of the bow-tie. Each along-track lateral buoyancy gradient is thus associated with a specific separation scale and a timestamp. Glider-based density gradients can be affected by internal waves. To filter the impact of internal waves on the PDF of horizontal buoyancy gradients, only gradients computed at a scale of twice the Rossby radius $\pm$ 1 km are considered. Rossby radii are estimated from the glider data and are $\sim8$ km in winter and $\sim20$ km in summer.

\subsection{Particle Tracking Experiments}

\subsubsection{Particle Advective Scheme}
To quantify the impact of submesoscale dynamics on the export of Particulate Organic Matter (POC), Lagrangian particle tracking is used in conjunction with the flow field obtained from the two experiments described above. Particles are tracked using the non-divergent velocity field at each face of a grid cell. The velocity field is then linearly interpolated onto the particle's position. For example, the eastward velocity (x-axis) of a particle is given by:

\begin{equation}
u(x) = u_{i-1} + \frac{(x - x_{i-1})}{(x_i - x_{i-1})}(u_{i}-u_{i-1})
\end{equation}
where the subscripts $i-1$ and $i$ denote the western and eastern walls of the grid cell where the particle is located, respectively. This can be re-written as:

\begin{equation}
\frac{\partial x}{\partial t} + \beta x + \delta =  0
\label{eq: eq_diff}
\end{equation}
where $\beta = (u_{i}-u_{i-1})/\Delta x$ and $\delta = -u_{i-1} - \beta x_{i-1}$ \citep{Doos_2013}. This differential equation can be solved analytically for $\beta \neq 0$:

\begin{equation}
x_{t_1} = \left(x_0 + \frac{\delta}{\beta}\right)\exp^{-\beta(t_1 - t_0)} - \frac{\delta}{\beta}
\label{eq: PT_advection}
\end{equation}
The time it will take for the particle to reach the eastern or western face of the grid cell can be computed by taking $x_{t_1} = x_i$ or $x_{t_1} = x_{i-1}$, respectively, and solving for $t_1$. For each advective timestep, the times required for the particle to reach any of the 6 walls of the grid cell are computed using (\ref{eq: PT_advection}). If any of those times is shorter than the advective timestep, the particle is advected until it reaches the cell wall. Then the flow field in the adjacent grid cell is considered and the particle is advected over the remaining time.

\subsubsection{Particle Seeding}

For all particle-tracking experiments, a single particle seeding event is prescribed. In the horizontal, particles are seeded every 250 m over the entire domain in the x-direction, and for 100 $<y<$ 200 km in the y-direction. The seeding is centered over the mean position of the middle front (see Figure \ref{fig: Papa_summer}) and is therefore not affected by undesired effects created by the solid north-south solid boundaries. In the vertical, particles are seeded every 1 m between 75 and 85 m. This depth range is chosen as it corresponds to the average euphotic depth at Station Papa, defined by the 1\% light level. The euphotic depth was computed for the months of February and June over the period 2007-2016 from profiles of Photosynthetically Available Radiation (PAR) collected at Station Papa as part of the long-term monitoring of Line P executed by the Department of Fisheries and Ocean Canada\footnote{https://www.waterproperties.ca/linep/index.php}. The average euphotic depth computed for both of these months is around 80 m, which agrees with previously established estimates of the euphotic depth \citep{Sherry_1999,Harrison_2004}.

In each particle-tracking experiment, four different classes of particles are released. Each particle class is characterized by a different sinking velocity: 0.025, 0.05, 1, and 5 m/day. The slowest-sinking class is essentially selected to represent non-sinking particles: based on the setup of our experiments, the slowest-sinking particles would take 800 days on average to be exported to a depth of 100 m through gravitational sinking, a timescale much greater than commonly observed remineralization timescales. The fastest-sinking velocity is chosen as a low end-member representing velocity classes that will be exported in its entirety over the course of our experiment. The results presented for the 5 m/day sinking class can therefore be extrapolated to any class with a higher sinking velocity.

The advective timestep for particles is set to 1.5 hours. The flow field is linearly interpolated in time between model outputs, justifying the higher temporal resolution used for particle tracking in \textit{Papa\_winter}. Particle positions are saved every 3 hours, along with key model variables interpolated onto the particle positions (e.g., density, vorticity). Particles are tracked for three weeks (28 days). Each particle-tracking experiment contains 1,971,717 particles per sinking-velocity class, for a total of 9,858,585 particles. Particles located deeper that the maximum winter mixed layer \citep[i.e., 100 m;][]{Pelland_2016, Plant_2016} are considered exported, as they will likely not be re-entrained into the mixed layer. %The export horizon is therefore set to 100 m, based on the seasonal cycle observed at Station Papa \citep{Pelland_2016}.

\subsubsection{Density and Biomass Spectra}
\label{sec: equations_biomass}
The slope of the size spectrum of particles is traditionally described by the slope of the log-log curve, $\xi$, between particle number $N$ and particle radius $r$ \citep[also know as the Junge slope;][]{White_2015}:

\begin{equation}
N(r) = N_0 \left(\frac{r}{r_0}\right)^{-\xi}
\label{eq: powerlaw_radius}
\end{equation}
where $N_0$ and $r_0$ represent a reference particle number and radius, chosen arbitrarily. For small particles ($<$400 $\mu$m) and relatively low temperature ($<$15$^o$C), it has been shown that the relationship between particle radius $r$ and sinking velocity $w_s$ exhibits a range of variation and is difficult to determine empirically. Nevertheless, Stokes' law, where $w_s \propto r^2$, is often used as a lower-bound sinking velocity estimate \citep{Bach_2012}. %\cite{Bach_2012} also demonstrate that Stokes law is generally a lower-bound estimate of sinking velocities for a specific particle size. 
Assuming a Stokes-like relationship, we can construct a particle sinking velocity spectrum $N(w_s)$ based on (\ref{eq: powerlaw_radius}), where $w_{s_0}$ is the sinking speed of particles with a radius $r_0$:

\begin{equation}
N(w_s) = N_0 \left(\frac{w_s}{w_{s_0}}\right)^{-\xi/2}
\label{eq: powerlaw_velocity}
\end{equation}
For a specific slope and sinking-velocity class, an equivalent number of particles per simulated particle can thus be computed using (\ref{eq: powerlaw_velocity}) (See Figure \ref{fig: sinking_velocity_spectrum}). For example, using the largest sinking velocity class as a reference (i,.e., $w_{s_0} = 5$ m/day and $N_0=$1,971,717), and a spectral slope $\xi = 4$, each simulated particle with a sinking velocity of 0.025 m/day in fact represent 40,000 particles (Figure \ref{fig: sinking_velocity_spectrum}). The relative biomass of a particle in a specific sinking-velocity class, $B_p(w_s)$, can be estimated if the biomass is assumed to scale with the particle's volume. The relative biomass of one particle in a sinking-velocity class $w_s$ can therefore be computed using:

% FIG4
\begin{figure}[ht]
	\centering
	\includegraphics[width = .49\linewidth]{figures/Fig4b.png}
	\includegraphics[width = .49\linewidth]{figures/Fig4a.png}
	\caption{Relative number of particles (left) and biomass (right) as a function of sinking velocity $w_s$. Sinking velocity spectrum are shown for three different Junge slope $\xi$: 2 (dotted), 3 (solid), and 2 (dashed). Colored squares indicate the sinking velocities of the three particle classes modeled: 0.025 m/day (red), 1 m/day (green), and 5 m/day (blue).}
	\label{fig: sinking_velocity_spectrum}
\end{figure}

\begin{equation}
	B_p(w_s) = B_p(w_{s_0})\left(\frac{w_s}{w_{s_0}}\right)^{3/2}
	\label{eq: biomass_realparticle}
\end{equation}
where $B_p(w_{s_0})$ is the biomass of a particle in the sinking velocity class $w_{s_0}$. The total biomass associated with one simulated particle can be obtained by scaling (\ref{eq: biomass_realparticle}) by the ratio $N(w_s)/N_0$:

\begin{equation}
	B(w_s) =  B_0\left(\frac{w_s}{w_{s_0}}\right)^{3/2}\frac{N(w_s)}{N_0}
	\label{eq: biomass_simulatedparticle}
\end{equation}
where $B_0 = B_p(w_{s_0})$. Combining (\ref{eq: powerlaw_velocity}) and (\ref{eq: biomass_simulatedparticle}) yields an expression relating the biomass associated with a simulated particle for a specific sinking-velocity class and the spectral slope (Figure \ref{fig: sinking_velocity_spectrum}):
\begin{equation}
	B(w_s) = B_0\left(\frac{w_s}{w_{s_0}}\right)^{\frac{3-\xi}{2}}
	\label{eq: biomassjungeslope}
\end{equation}
Using the same example as before where $\xi = 4$, if the amount of biomass associated with one simulated particle in the 5 m/day sinking-velocity class is taken as $B_0 = 1$, then one simulated particle sinking at 0.025 m/day contains 14.14 units of biomass and a single particle contains $14.14/40,000 =3.5\times10^{-4}$ units of biomass (see Figure \ref{fig: sinking_velocity_spectrum}). This normalized formulation of particle number and biomass (see Equations (\ref{eq: powerlaw_velocity}) and (\ref{eq: biomassjungeslope})) has the advantage that the impact of spectral slope on the relative export of biomass can be quantified without needing a large number of particle-tracking experiments, where the number of seeded particles would vary to account for the different spectral slopes. For the purpose of this study, only the relative amount of biomass is relevant. For simplicity, we define a normalized biomass unit for $\xi =3$ as $B_0=$ 1. The values taken by $B_0$ for other Junge slopes $\xi$ are computed under the condition that the total amount of biomass is kept constant (Figure \ref{fig: sinking_velocity_spectrum}b).

\subsubsection{Particle Remineralization Scheme}
\label{sec: remin_scheme}
Remineralization of particles as they sink through the water column impacts the amount of biomass exported. Slow-sinking particles generally contains less biomass and spends more time in the mixed layer, which means that they are remineralized at a shallower depth than faster sinking particles. Remineralization processes are complex, species-dependent, and generally not well-understood. In the absence of a consensus on a general functional form of particle remineralization, we rely on an idealized relationship which assumes that the biomass content of a particle decreases in time proportionally to the particle volume. Remineralization is thus modeled as an exponential decrease of biomass with time at a rate $k$ \citep{Iversen_2010, Iversen_2013}:
\begin{equation}
	B (t) = B^0\exp(-kt)
	\label{eq: remin_scheme}
\end{equation}
where $B^0$ denotes the biomass content at $t = 0$ days, and the remineralization rate is taken to be $k = 0.13$ day$^{-1}$ in this study \citep{Iversen_2010}. This remineralization rate is independent of particle sinking velocity, and seems to lie within the range of other estimates \citep{Ploug_2008, Iversen_2010, Iversen_2013}. The change in biomass with time is in turn expected to affect the sinking velocity of the particle. Given that $B \propto w^{3/2}$ (see Equation (\ref{eq: biomass_realparticle})), particles in all sinking-velocity classes undergo a decay in sinking speed according to:

\begin{equation}
w_s (t) = w_s^0\exp(-\frac{2kt}{3})
\label{eq: sinking_velocity_remin}
\end{equation}
where $w_s^0$ is the initial sinking velocity at $t = 0$ days. In this study, the impact of remineralization is thus considered through the implementation of a time-dependent sinking velocity (Equation \ref{eq: sinking_velocity_remin}). While particles classes are classified based on their initial sinking-velocity, it is worth noting that over the length of the particle-tracking experiments that include remineralization (28 days), particle sinking speeds slow down to 10\% of their initial velocity.

\section{Results}
\label{sec: Results}
\subsection{Seasonally varying dynamical regimes}

Two model experiments are designed to capture different dynamical conditions observed in the Northeast Pacific Ocean in summer and winter. \textit{Papa\_summer} is initialized in early spring (doy 105) when the water column is characterized by a relatively deep mixed layer ($\sim$100 m) and a halocline located between 100 and 150 m (Figure \ref{fig: Papa_summer}). The forcing by a realistic, positive, net heat flux generates the restratification of the water column, with the development of a strong thermocline between 25 and 50~m leading to the shoaling of the mixed layer and a subsurface peak in $N^2$ at about 30 m (see Figure \ref{fig: Papa_summer}). A comparison between model outputs and monthly-averaged density profiles from underwater gliders collected in June and July over the period 2008-2009 yields correlation coefficients of $r = 0.87$ and $r = 0.88$, respectively. These high correlation suggest that \textit{Papa\_summer} numerical experiment captures the vertical spring and summer conditions in the Northeast Pacific Ocean.

In the horizontal, the prescribed density fronts progressively become unstable within the first 60 days of the experiment (Figure \ref{fig: Papa_summer}). During this time, the Total Kinetic Energy (KE$_{\text{tot}}$) contained in the model domain slowly increases before reaching a maximum at doy 162, where it remains relatively constant for the rest of the simulation. The flattening of the KE$_{\text{tot}}$ curve is used to determine the time necessary for the simulation to spin-up, hence determining the start day of the particle-tracking experiments. The ocean dynamics associated with \textit{Papa\_summer} are characterized using PDFs of horizontal buoyancy gradients ($M^2 = |\nabla_Hb|^2$), vertical velocities ($w$), and Rossby numbers computed from the normalized vertical component of the relative vorticity (Ro $=(v_x - u_y)/f$ where f = 1.12$\times10^{-4}$; Figure \ref{fig: dynamics}).

%FIG5
\begin{figure}[t]
	\centering
	\includegraphics[width = 1\linewidth]{figures/Fig5_dynamic_signature_w_valid}
	\caption{Snapshots of $M^2$ (top), $\zeta /f$ (middle), and $w$ (bottom) half-way through the particle tracking experiment for \textit{Papa\_summer} (left) and \textit{Papa\_winter} (right), with the Mixed Layer Depth indicated by the solid black line. The corresponding Probability Distribution Functions (PDFs) are shown in the center for both \textit{Papa\_summer} (blue) and \textit{Papa\_winter} (red). Note the different colorbars used for \textit{Papa\_summer} and \textit{Papa\_winter}. Histograms of $M^2$ computed from glider data at Station Papa in February (blue line) and July (red line) are superimposed in the top middle panel.}
	\label{fig: dynamics}
\end{figure}

%The largest horizontal buoyancy gradients are observed above the thermocline ($z$$<$50 m) and along the halocline (100--200 m), and range between 0 and 1.2$\times 10^{-7}$ s$^{-2}$ (Figure \ref{fig: dynamics}). 
Lateral buoyancy gradients in the summer are relatively weak $\mathcal{O}$(10$^{-8}$ s$^{-2}$) and result in low Rossby numbers $\mathcal{O}$(0.1), with positive relative vorticity on the denser (north) side of the front and negative relative vorticity on the lighter (south) side of the front. Corresponding vertical velocities are consistently weaker than 1 m/day ($<$10$^{-5}$ m/s) and are characterized by regions of weak upwelling and downwelling on 10 km scales, associated with the meandering of the front \citep{Bower_1989}. Alternating bands of upwelling and downwelling at $\mathcal{O}$(1 km) spatial scale are superimposed, and likely caused by propagating internal waves. Coherent vertical velocities structures extend to depths much greater than the mixed layer depth ($\sim$25 m; Figure \ref{fig: dynamics}). The amplitude of the vertical velocity field coincides with the expected order of magnitude given by the scaling $w \propto RofU/N$ \citep{Mahadevan_2016}: using $Ro\sim0.1$ (Figure \ref{fig: dynamics}), $N\sim10^{-2}$ s$^{-1}$ (Figure \ref{fig: Papa_summer}), $f\sim10^{-5}$ s$^{-1}$, and $U\sim0.01$ m/s, we obtain $w\sim 10^{-6}$ m/s, or $\sim 10^{-1}$ m/day. %The agreement between the scaling and the model outputs suggest that PSOM is capable of reproducing a realistic vertical flow field.

\textit{Papa\_winter} is, on the other hand, initialized in the winter (doy 0) to capture a time period where the mixed layer depth is deeper ($\sim$ 100 m) and density gradients more pronounced \citep{Pelland_2016}. At this time of year, the water column in this region is characterized by the presence of a deep halocline between 100 and 150 m \citep[Figure \ref{fig: Papa_winter}][]{Pelland_2016}. After spin-up, the vertical stratification remains consistent throughout the model run, and compares well with the vertical profile obtained from glider observations for the month of March ($r = 0.95$; see Figure \ref{fig: Papa_winter}). In the horizontal, prescribed density fronts are much sharper than in summer (i.e., over smaller spatial scales O(1 km ) vs. O(10 km)). Because of these stronger density gradients, combined with the alternating zonal winds and constantly negative surface heat flux, the fronts become unstable more rapidly than in summer (Figure \ref{fig: Papa_winter}). As a result, KE$_{\text{tot}}$ starts to plateau at doy 48. The experiment is considered spun-up by doy 50 and the particle-tracking experiment is initialized. %Many sharp fronts and filaments on the order of 1 km can be seen in \textit{Papa\_winter}, penetrating to ocean upper layer to the deeper than in the summer, dissipating within the mixed layer at $\sim$100 m (Figure \ref{fig: dynamics}).

The frontal structures visible in the horizontal buoyancy gradient field are associated with filaments of relatively high Rossby number of $\mathcal{O}$(1) (Figure \ref{fig: dynamics}). The PDF of relative vorticity reveals a positively-skewed distribution ($s = 0.68$). This is in agreement with the fact the relative vorticity is more likely to be cyclonic than anticyclonic, based on conservation of potential vorticity \citep{Hoskins_1972}. Regions with a high Rossby number are localized and located in the mixed layer exclusively. In places where the local Rossby number reaches $\mathcal{O}$(1), geostrophic balance is lost and a vertical secondary ageostrophic circulation begins to slump the isopycnals and restore the flow to a more geostrophically-balanced flow. This ageostrophic secondary circulation therefore generates ``hot spots" of higher vertical velocities. The fine-scale structures in the vertical velocity field corresponding to $\mathcal{O}$(1) Rossby numbers  can be seen in Figure \ref{fig: dynamics}, with local vertical velocities up to 60 m/day ($\sim 7\times10^{-4}$ m/s). Contrary to the PDF of relative vorticity, the distribution of vertical velocities demonstrate a negative skewness ($s =$ -0.25). This is in agreement with the theory: In fact, positive relative vorticity is associated with the dense side of a density front, where vertical velocities are negative \citep{Mahadevan_2016}. Once again, the amplitude of these vertical velocity hot spots is coherent with the scaling $w \propto RofU/N$: using $Ro\sim1$, $N\sim10^{-2}$ 1/s, $f\sim10^{-5}$ 1/s, and $U\sim0.1$ m/s, we obtain $w\sim 10^{-4}$ m/s, or $\sim 10^{1}$ m/day. %This demonstrates the capabilities of PSOM to resolve sharp frontal features at the kilometer scale, as well as the ageostrophic circulation associated with those fronts.

Comparing \textit{Papa\_summer} and \textit{Papa\_winter} highlights the different dynamical regimes in the two experiments. In \textit{Papa\_winter}, density fronts tend to be sharper, meaning larger density gradients over shorter spatial scales. When computed at the kilometer-scale, the PDF of horizontal buoyancy gradients in \textit{Papa\_winter} exhibits a longer tail than in \textit{Papa\_summer} (Figure \ref{fig: dynamics}). When compared to observations, the PDFs of $M^2$ in \textit{Papa\_summer} and \textit{Papa\_winter} demonstrate a correlation with observations of $r = 0.93$ and $r = 0.95$, respectively

The wider PDF of vertical velocities in \textit{Papa\_winter} shows advective velocities that match and exceed typical gravitational sinking velocities, particularly for smaller, and therefore slower-sinking, particulate organic material. The secondary ageostrophic circulation that develops at submeso-scales (i.e., Ro O(1)) therefore generates an export mechanism that directly competes with the traditional paradigm that relies on gravitational sinking leading the export of particulate matter in the ocean.

\subsection{Gravitational and Advective Sinking of POC}

Both model experiments described above were then used to investigate the relationship between ocean dynamics and particle downward flux, using Lagrangian particle-tracking. Domain-averaged, downward particle flux is expected to be a combination of the flux driving by gravitational sinking ($\left<w_sB\right>$), and by the vertical advective currents affecting the particle along its pathway ($\left<wB\right>$). The deviation in particle depths from the traditional one-dimensional gravitationally driven model is shown in Figure \ref{fig: particle_depth_distribution}d for both summer and winter cases. In the summer, the PDF of particle density versus depth remains relatively narrow through time, and is centered on a depth level that can be predicted using a simple 1D gravitational model (see shaded curves in Figure \ref{fig: particle_depth_distribution}). The spread in the particle density also vary little among particle classes with different sinking velocities, suggesting that downward fluxes of particles is greatly dominated by gravitational settling and is not subject to significant vertical ocean currents.

%FIG6
\begin{figure}[ht!]
	\centering
	\includegraphics[width = .8\linewidth]{figures/Fig6.png}
	\caption{[left] Gridded anomaly in the median depth of particles 25 days after particle release, in the Winter experiment. Blue (red) grid cells indicate that the median depth of particles in this cell is deeper (shallower) than expected based on a 1D gravitational model where $z = w_s\timest$. [right] Probability Distribution Function (PDF) of particles as a function of depth for each velocity class. The winter distribution is shown as thick lines, while the summer distribution is represented by the shaded regions. Triangle markers indicate the expected depth of particles after 25 days based on the 1D gravitational model, which is used as a reference to compute the depth anomalies. Release depth is indicated by the thick dashed line.}	
	\label{fig: particle_depth_distribution}
\end{figure}

In the winter, however, PDFs of particle density versus depth is wider, in agreement with the stronger vertical ocean currents occurring in the winter (see Figure \ref{fig: dynamics}). A top-view of the deviation in the downward particle flux from the traditionally considered 1D gravitational model can be seen in Figure \ref{fig: particle_depth_distribution} (panels (a)-(c)). Slower-sinking particles deviate more than faster-sinking particles, exhibiting median depth anomalies up to 50 m. This is due to the fact that slower-sinking particles spend more time in the mixed layer, where most of the stronger vertical currents tend to occur (Figure \ref{fig: dynamics}). An interesting result emerges from the spatial distribution of the depth-anomaly: both positive (i.e., particles are shallower than expected) and negative (i.e., particles are deeper than expected) anomalies are organized into features with a length-scale $\mathcal{O}$(1-10 km). This further highlights the importance of winter submesoscale circulation for vertical fluxes of particles.%, which is explicitly shown in Figure \ref{fig: biomass_export}.

A relative amount of biomass is associated to the particles using Equation \ref{eq: biomassjungeslope}. PDFs of relative biomass as a function of the vertical velocity is shown in Figure \ref{fig: biomass_export}. Following the traditional paradigm derived from the simple 1D gravitational model, the downward flux of biomass in the summer is dominated by faster-sinking particle classes capable of carrying particulate material downwards more efficiently. The contribution of slower-sinking particles, however, critically depends on the slope of the size spectrum (see Figure \ref{fig: sinking_velocity_spectrum}). As the Junge slope increases, the spectrum of biomass steepens, and the relative contribution of slower-sinking particles to the downward biomass flux significantly increases. In fact, the contribution of slower-sinking particles to the summer downward flux increases by a factor 100 (from 0.1\% to 10\%) when the Junge slope varies from $\xi =2$ to $\xi = 4$. While significant, the impact of a change in the Junge slope in summer conditions does not challenge the dominant role played by faster-sinking particles. This result can be explained by the fact that, in the summer, vertical velocities are weak and vertical biomass fluxes are therefore gravitationally-driven ($\left<w_sB\right> > \left<wB\right>$).

%FIG7
\begin{figure}[t]
	\centering
	\includegraphics[width = 1\linewidth]{figures/Fig7.png}
	\caption{Probability Distribution Function (PDF) of relative biomass versus vertical velocity along particle trajectories in the summer case [left] and winter case [right], with a Junge slope of 2 [top] and 4 [bottom]. Inserts show the integrated relative downward biomass flux associated with each sinking-velocity class. Both winter dynamics and steeper Junge slopes tend to increase the relative contribution of slower-sinking particles.}
	\label{fig: biomass_export}
\end{figure}

In the winter, PDFs of relative biomass as a function of vertical velocities present a much larger spread, with velocity magnitudes exceeding 50 m/day. For $\xi = 2$, the relative contribution of slower-sinking particles to the downward flux significantly increases from 0.1\% in the summer to about 4\% in the winter, demonstrating the impact advective velocities alone can have on vertical fluxes. Nevertheless, slower-sinking particles remain a relatively small contributor to the total downward flux of biomass (Figure \ref{fig: biomass_export}b). When winter ocean dynamics are coupled with a steeper Junge slope, however, slower-sinking particles largely dominate the downward biomass flux. In our winter simulations with $\xi = 4$, we find that the slowest-sinking particle class is responsible for about 80\% of the biomass flux.

Our results show that both a steepening of the particle size spectra and the presence of submesoscale dynamics can enhance the contribution of slower-sinking particles to the downward biomass flux. While the former is simply due to an increase in particle density in slower-sinking particle classes, the latter is attributed to the larger vertical velocity generated by submesoscale instabilities. When both are combined, as expected in the wintertime, slower-sinking particles then become the leading contributor to the downward biomass transport. However, slower-sinking particles are generally expected to remineralize on timescales shorter than their export timescale, fueling the argument that the focus should be upon faster-sinking particle classes. The impacts of remineralization on export are thus considered in the following section to test the robustness of the findings.

 \subsection{Particle Remineralization}
 \label{sec: results_remin}

 Both submesoscale dynamics and the Junge slope were identified as key factors impacting the respective role played by different particle classes in driving downward biomass fluxes. Simple Lagrangian particles were used to isolate the effects of these two factors. In reality, however, sinking velocities of particulate matter varies in time as the particles slowly remineralize. A remineralizing behavior was therefore implemented for the Lagrangian particles, using Equation \ref{eq: sinking_velocity_remin}, to investigate the impact that remineralization processes have on our findings. The traditional paradigm relies on the fact that slow-sinking particles tend to fully remineralize over short timescales, further enhancing the importance of faster-sinking particles classes in driving downward biomass fluxes. While this paradigm holds for flatter Junge slope, where the biomass content is dominated by faster-sinking particles, it becomes unfit at steeper slopes.

 Figure \ref{fig: biomass_export_remin} compares the relative biomass and downward biomass fluxes associated with each of the modeled particle classes for $\xi = 4$ with and without the remineralization scheme. As previously detailed, downward fluxes of biomass are dominated by faster-sinking particles during summertime and in the absence of remineralization. This is due to the fact that the flux of biomass $\left<w_{tot}B\right> = \left<w_sB\right> + \left<wB\right>$ is driven by $\left<w_sB\right>$, despite a smaller relative biomass content per particle. This is characteristic of a gravitationally-driven system, where settling velocity dictates the contribution to downward fluxes. Implementing remineralization processes, however, directly affects the particle settling velocity which slows down as particles remineralize. This effect can be seen in Figure \ref{fig: biomass_export_remin}c, where PDFs of relative biomass per particle class are shifted towards weaker vertical velocities than in the absence of remineralization, as predicted by Equation \ref{eq: sinking_velocity_remin}. As a result, the gravitationally-driven term $\left<w_sB\right>$ decreases with time, and the downward flux of biomass becomes generally advectively-driven by day 25 (Figure \ref{fig: biomass_export_remin}). 
   %FIG8
 \begin{figure}[t!]
	\centering
	\includegraphics[width = 1\linewidth]{figures/Fig8.png}
	\caption{Probability Distribution Function (PDF) of relative biomass versus vertical velocity along particle trajectories in the summer case [left] and winter case [right], ignoring [top] and including [bottom] particle remineralization, for $\xi$ = 4 and 25 days after particle release. Inserts show the integrated relative downward biomass flux associated with each sinking-velocity class. Remineralization processes have the greatest impact on fast-sinking particles, especially in summer dynamics.}	
	\label{fig: biomass_export_remin}
\end{figure}
In a advectively-driven system where $\left<w_sB\right> < \left<w_B\right>$, the relative amount of biomass content in a particle class becomes important and dictates the respective contribution of each particle class to the total downward biomass fluxes. This shift from a gravitationally-driven to an advectively-driven system is observed when implementing particle remineralization in the summer (Figure \ref{fig: biomass_export_remin}c): in the absence of remineralization, faster-sinking particles dominate the downward biomass fluxes (60\%; see Figure \ref{fig: biomass_export_remin}a). When remineralization processes are considered, slower-sinking particles become the dominant contributor to biomass fluxes (see Figure \ref{fig: biomass_export_remin}c). As shown in Figure \ref{fig: biomass_export}, downward biomass fluxes in the wintertime are generally advectively-driven, due to the larger vertical velocities associated with wintertime ocean dynamics (82\%; see Figure \ref{fig: biomass_export_remin}b). Even after implementing our remineralization scheme, slower-sinking particles remain the largest contributor to downward biomass fluxes (87\%; see Figure \ref{fig: biomass_export_remin}d).

 These results highlight the importance in considering slower-sinking particle classes when considering downward biomass fluxes. It also demonstrates that, contrarily to the traditional paradigm, remineralization processes enhance the role played by slower-sinking particles in biomass fluxes, in cases where the biomass spectrum slope is negative.

 The timescales over which the system transitioned from an gravitationally-driven to an advectively-driven system depends on the remineralization model used.

\section{Discussion}
\label{sec: Discussion}

\subsection{Dynamical Regimes}
\label{sec: Discussion_model}
\textit{Papa\_summer} and \textit{Papa\_winter} experiments were designed to statistically capture the ocean dynamics at Station Papa (145$^o$W, 50$^o$N) in the Northeast Pacific Ocean. After spin-up, the model demonstrated similar distribution of both horizontal ($M^2$) and vertical ($N^2$) density gradients to observations from underwater gliders (see Figures \ref{fig: Papa_summer}, \ref{fig: Papa_winter}, and \ref{fig: dynamics}). The two experiments, however, show significantly different distributions of $M^2$, with the winter distribution exhibiting a longer tail, suggesting sharper density gradients. The tail of the wintertime distribution is only partially captured by the glider data, due to the fact that underwater gliders sample gradients at a maximum resolution of about 10 km, while the model resolution is 500 m and therefore allows sharp submesoscale fronts and filaments to be captured in the distribution.

Studies investigating submesoscale dynamics traditionally focused on regions where the presence of submesoscale fronts and filaments are established, such as western boundary currents with strong gradients \citep{Dasaro_2011, Thomas_2013}, or at the edge of mesoscale features \citep{vanHaren_2006,Waite_2016}. The seasonality in submesoscale dynamics captured in the glider dataset at Station Papa, and reflected in the model experiments, echoes findings of recent observational studies conducted at a similar latitude in the Atlantic Ocean demonstrating the intensification of submesoscale dynamics in the wintertime \citep{Thompson_2016, Buckingham_2016}. Despite being sometimes qualified as an ``eddy desert" with low kinetic energy \citep{Chelton_2011}, ocean characteristics in the eastern part of the Pacific subpolar gyre suggest the presence of submesoscale features in the wintertime: strong density gradients, localized Rossby numbers of order 1, a balanced Richardson number $Ri_b = \frac{f^2N^2}{M^4}$ smaller than 1, a positively skewed distribution in vorticity, and a negatively skewed distribution of vertical velocities \citep[see Figure \ref{fig: dynamics};][]{Thomas_2008,Rudnick_2001, Buckingham_2016}.

Strong downward velocities are hypothesized to enhance POC export by advecting slower-sinking particles out of the mixed layer. \textit{Papa\_winter} indeed exhibits vertical velocities more than 20 times larger than in \textit{Papa\_summer}. The vertical currents in \textit{Papa\_winter}, however, tend to be much patchier than the weaker vertical currents observed in \textit{Papa\_summer}. Because both particle production and downward vertical velocities present a high degree of patchiness, it requires a certain level of covariance between the two fields for the export to effectively be enhanced \citep{Mahadevan_2012}. A more realistic seeding strategy for Lagrangian particles, such as one guided by biological tracers, would likely provide important information towards a better understanding of the effects of patchiness on POC export at submeso-scales

The hypothesis tested in this study is that submesoscale activity enhances export of particulate matter at Station Papa by shortening the export timescale of particulate matter. The wintertime intensification in submesoscale activity has the potential to indeed enhance export (see discussion in Section \ref{sec: discussion_particle}). However, the seasonal cycle in submesoscale activity is out of phase with the one in net community productivity, which peaks in the spring and summertime when the mixed layer is shallower \citep{Plant_2016}. Two mechanisms are therefore present to potentially sustain a year-long POC export flux: In the winter, less particulate material is present in the mixed layer, but active submesoscale dynamics tend to enhance the POC export flux by advecting the more numerous slower-sinking particles into the ocean interior. In the summer, the production of POC is at its yearly maximum, but export tends to be dominated by gravitational sinking, which favors faster-sinking particles and thus exclude part of the particle spectrum from contributing to the export flux.

\subsection{Downward Fluxes}
\label{sec: discussion_particle}

Analyses of particle tracking experiments reveal that the role played by slower-sinking particles in downward particulate fluxes depends on two main characteristics: (1) the dynamics of the ocean flow field, and (2) the slope of the size spectrum (i.e., the Junge slope $\xi$).

Mixed layer ocean dynamics at station Papa change significantly between the winter and the summer. In the winter, submesoscale dynamics are intensified, and sharp fronts and filaments develop in the mixed layer. This seasonal change in dynamics is consistent with recent observations \citep{Thompson_2016, Buckingham_2016}, and models \citep{Brannigan_2015, Callies_2015, Rocha_2016}, characterizing the seasonal cycle of submesoscale dynamics. The winter intensification in submesoscale dynamics was proven to have an important impact on the downward flux of all sinking-velocity classes modeled in this experiment.

In the summer, the gravitationally-driven dynamics efficiently drives a downward flux dominated by faster-sinking particles, less numerous but generally richer in POC, and little to no slower-sinking particles. In the winter, however, vertical fluxes tend to be advectively-driven, which leads to a slightly weaker downward flux of faster-sinking particles than in the summer due to resuspension, but a much larger flux of slower-sinking particles, which are present in far greater numbers (Figure \ref{fig: biomass_export}). These two seasonal mechanisms raise the question as to which process is most efficient in driving a downward flux of particulate material.

In the absence of remineralization, both a steeper size spectrum slope ($\xi>3$ in this case) and enhanced submesoscale dynamics increase the contribution of slower-sinking particle classes to the downward biomass flux. This is only when both of these conditions are combined, however, that slower-sinking particles dominate the downward flux of biomass (Figure \ref{fig: biomass_export}). This is a significant result, as Junge slopes greater than 3 have been observed in the ocean \citep{Kostadinov_2009, White_2015}. The threshold value of $\xi =3$ for a change in the biomass spectral slope (see Figure \ref{fig: sinking_velocity_spectrum}b) is of course a consequence of first-order approximations used in this study describing the relationships between particle size, sinking velocity, and biomass content. Nevertheless, our results demonstrate the importance of resolving the lower range of the particle spectrum when measuring vertical fluxes, especially when submesoscale dynamics are active. It also highlights the importance of better constraining the relationships linking particle size, sinking velocity, and biomass content.

Introducing remineralization processes significantly decreases the biomass flux. Counter-intuitively, however, the implementation of a remineralization scheme further strengthen the role played by slower-sinking particles in the biomass flux (Figure \ref{fig: biomass_export_remin}. This can be explained by the fact that remineralization processes have a greater impact on sinking-velocity classes that rely on gravitational sinking to be exported, as these particles decelerate as they remineralize. In the summer, all particle classes are similarly affected by the remineralization scheme as downward fluxes are gravitationally-driven. In the winter, however, slower-sinking particles are exported through advective processes. Their export timescale is little affected by remineralization processes as it only depends on local ocean dynamics.

The result of this study suggest that slow- and non-sinking particles must be considered when studying the downward flux of particulate matter in the upper ocean. The patchiness associated with both particle production and submesoscale features poses a real observational challenge to properly resolve vertical fluxes. Based on our findings, subsequent studies should focus on testing the impact of patchiness on vertical fluxes: In the wintertime, when size spectral slope is steep and submesoscale dynamics most active, vertical fluxes could be grossly underestimated depending on the level of co-occurrence between particle production and stronger vertical currents.

\section{Conclusion}
\label{sec: Conclusions}
The main conclusions of this study are: 
\begin{enumerate}
\item Ocean dynamics in the subpolar Northeast Pacific exhibit a seasonal cycle with low submesoscale activity in the summertime, and more submesoscale features present in the wintertime. Submesoscale dynamics generate larger, and asymmetrical, vertical currents leading to a vertical biomass flux driven by advective processes, as opposed to gravitationally-driven in the summertime.

\item Submesoscale dynamics generally enhance downward fluxes by increasing the contribution of slower-sinking particles to the total flux by increasing the number of particle advected downward by ocean currents. The contribution of slower-sinking particles to the total flux is found to be at least significant, and can be dominant in certain conditions.
\item Contribution of slower-sinking particles to downward biomass fluxes export depends on the slope of the particle size spectra (i.e., Junge Slope), that controls the relative number of particles per size class. Two cases emerge from this study:
\begin{enumerate}
	\item If the Junge slope is smaller than 3, larger particles contribute most to vertical biomass fluxes independently of flow dynamics, as there are no mechanism capable of selectively advecting slower-sinking particles. The system is described as gravitationally-driven.
	\item If the Junge slope is greater than 3, as most commonly observed, ocean dynamics become the determining factor on which particle classes dominate the downward fluxes: as submesoscale dynamics become more active, an ageostrophic circulation leading to larger vertical velocities develops. In these conditions, downward biomass fluxes are largely driven by the slower-sinking particle classes.
	\end{enumerate}
\item Remineralization processes logically reduce the amplitude of biomass fluxes. However, it unexpectedly enhances the role played by slower-sinking particles as the impact of remineralization is larger on faster-sinking particles by affecting both biomass content and sinking velocity.
\end{enumerate}

\section*{Acknowledgments}

The authors acknowledge support from the National Aeronautics and Space Administration as part of the EXport Processes in the global Ocean from RemoTe Sensing (EXPORTS) program - grant NNX16AR48G. We would like to thank A. Thompson (Caltech) for his useful comments and suggestions, N. Pelland (NOAA) and Charlie Eriksen (University of Washington) for sharing the glider data collected at Ocean Station Papa, as well as Sebastian Essink for his assistance in developing the particle-tracking code. All data used in this manuscript are publicly available: Ocean Station Papa data is available on PMEL's website (\url{https://www.pmel.noaa.gov/ocs/Papa}) and gridded Argo products can be downloaded at \url{http://www.seanoe.org/data/00348/45945}. Glider data is archived at NCEI: https://data.nodc.noaa.gov/cgi-bin/iso?id=gov.noaa.nodc:0155879. Model outputs, and particle trajectories are freely available on request. % Finally, we are grateful to the two anonymous reviewers for their valuable comments and suggestions.

%% ------------------------------------------------------------------------ %%
%%  REFERENCE LIST AND TEXT CITATIONS
%\bibliographystyle{apalike}
\bibliography{sinking_biblio}

%% ------------------------------------------------------------------------ %%
%
%  END ARTICLE
%
%% ------------------------------------------------------------------------ %%

%\end{article}
%\clearpage

%% ------------------------------------------------------------------------ %%
%
%  FIGURES AND TABLES
%
%% ------------------------------------------------------------------------ %%

\end{document}